Viscous Erosion in Porous Media

### 2018-01-29 docs/methods/erosion_methods.tex
NM
worked on intro

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion is one example of a fluid-mechanical process that occurs when a
fluid alters the geometry.  Erosion is responsible for a variety of
processes, many of which are geophysical.  For example, sand dunes are
longer on the windward side because of the wind erosion.  Another
applications is erosion caused by groundwater flow, where a slow moving
fluid erodes channels, or conduits, in the earth which creates a
heterogeneous media that has a large effect on the resulting flow.
Erosion is also important in the biophysical community. Plaque erosion
in arteries is key to understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}. In all these examples,
there is a feedback process between the flow and the erodible
bodies---the shape of the eroding body affects the flow, while the flow
changes the shape by fluidizing the eroding body.  This fluid-mechanical
feedback system has been simulated in idealized cases such as a single
eroding two-dimensional body in a uniform
flow~\cite{moo-ris-chi-zha-she2013} or a single mobile eroding
three-dimensional body~\cite{mit-spa2016}. To further develop insight
into the effect of two-dimensional erosion, we consider erosion of a
multiply-connected domain (Figure~\ref{fig:50bodies}), such as a porous
media that is typically used to model groundwater flow.

Because of the scales involved with groundwater flow, we will restrict
our attention to Stokesian dynamics.  For groundwater flow, Darcy's
equation are often used, but this assumes a homogeneous geometry which
is not always a fair assumption.  Instead, we will use the
incompressible Stokes equations to represent the flow, and the
individual pores will be be resolved, resulting in a complex
multiply-connected geometry.  A boundary integral equation is
well-suited as it can resolve complex geometries and the non-negligible
interactions between bodies are resolved, all the while high-order
accuracy and computational efficiency are achieved. To link the body and
fluid dynamics, the body is eroded in the normal direction at a rate
that is proportional to the magnitude of the shear
stress~\cite{ris-moo-chi-she-zha2012}. Other models for erosion include
ones that use an erosion rate that is proportional to a power of the
magnitude of the shear stress~\cite{par-izu2000}, or ones that use a
critical shear stress below which erosion does not occur~\cite{lag2000}.

@@

@@

% Pozrikidis page 239: symmetric flow.
Certain features of this shape can be predicted analytically by appealing to the principal of uniform stress distribution. In particular, consider the angle formed at the front and rear stagnation points. In long time, the stress becomes uniform, in particular in a small vicinity of this front angle in which body curvature can be ignored. Zooming in on the front angle, we can consider the problem of Stokes flow around an infinite wedge of opening angle $\oangle$. Exact solutions can be obtained to this problem via separation of variables in polar coordinates $(r, \theta)$ \cite{poz1997}. In this section alone, $\theta$ denotes the polar angle and not the tangent angle of the body.
The corresponding stream function is $\psi = r^{\lambda}f(\theta)$, where $f$ satisfies the 4th order ODE
\begin{equation}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0
\end{equation}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$, where $\thb = \pi - \oangle/2$, along with the condition of odd-symmetry $f(0) = f''(0) = 0$ (corresponding to up-down symmetry of the flow).
The corresponding shear stress evaluated on the surface of the wedge is $\tau = \mu r^{\lambda-2} f''(\thb)$. Thus, the wedge of uniform stress corresponds to $\lambda = 2$. For this value, the solution for $f$ satisfying odd symmetry is
\begin{equation}
f(\theta) = B \sin (2 \theta) + C \theta
\end{equation}
Then the eigenvalue problem

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Erosion is a fluid-mechanical process that is responsible for a variety
of geophysical and other processes.  For example, wind erosion causes
sand dunes to be  longer on the windward side~\cite{han1969}, and
erosion is responsible for branching structures of streams and
rivers~\cite{coh-dev-sey-yi-szy-rot2015}.  Erosion is also important in
the biophysical community. Plaque erosion in arteries is key to
understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}.  Another applications is
erosion caused by groundwater flow, where a slow moving fluid erodes
channels, or conduits, in the ground.  This results in a heterogeneous
media that has a large effect on the resulting flow.  In all these
examples, there is a feedback process between the flow and the erodible
bodies---the shape of the eroding body affects the flow, while the flow
changes the shape by fluidizing the eroding body.  This fluid-mechanical
feedback system has been simulated in idealized cases such as a single
eroding two-dimensional body in a uniform
flow~\cite{moo-ris-chi-zha-she2013} or a single mobile eroding
three-dimensional body~\cite{mit-spa2016}. To further develop insight
into the effect of two-dimensional erosion, we consider erosion of a
multiply-connected domain (Figure~\ref{fig:50bodies}), such as the
porous geometry of many groundwater flows.

Because of the scales involved with groundwater flow, we will restrict
our attention to Stokesian dynamics.  For groundwater flow, Darcy's law
is often used, but this assumes a homogeneous geometry which is often
not valid, and the method does not resolve the pore structure.  Instead, we
will use the incompressible Stokes equations to represent the flow, and
the individual pores will be be resolved, resulting in a complex
multiply-connected geometry.  A boundary integral equation is
well-suited as it can resolve complex geometries and the non-negligible
interactions between bodies are resolved. Moreover, high-order
accuracy and computational efficiency are obtained. To link the body and
fluid dynamics, the body is eroded in the normal direction at a rate
that is proportional to the magnitude of the shear
stress~\cite{ris-moo-chi-she-zha2012}.  Tracking the interface of
fluid-structure interaction problems with a long time horizon is
particularly---the mesh is easily distorted and tangled by the dynamics
and regions of high curvature develop.  We adopt a \thL~framework that
keeps the mesh equidistributed in arclength, and a curvature
penalization term can be added without introducing any non-linear
stiffness.
\todo[inline]{rework this}

@@

Several groups consider the optimal body in a Stokes fluid that
optimizes some quantity.  For example, Pironneau~\cite{pir1973} found
the axisymmetric shape of a fixed volume body that minimizes drag is
similar to an American football with an opening angle of 60 degrees, and
this shape corresponds to a body with constant surface vorticity.  This
analysis is taken further by Montenegro-Johnson and
Lauga~\cite{mon-lau2015} where the optimal axisymmetric shape of fixed
surface area is computed.  We include a similar calculation for the
two-dimensional Stokes equations, and the resulting shape agrees with
our numerical solution.

\todo[inline]{Dump of text}
Other models for erosion include ones that use an erosion rate that is
proportional to a power of the magnitude of the shear
stress~\cite{par-izu2000}, or ones that use a critical shear stress
below which erosion does not occur~\cite{lag2000}.
\todo[inline]{Dump of text}

@@

% Pozrikidis page 239: symmetric flow.
Certain features of this shape can be predicted analytically by appealing to the principal of uniform stress distribution. In particular, consider the angle formed at the front and rear stagnation points. In long time, the stress becomes uniform, in particular in a small vicinity of this front angle in which body curvature can be ignored. Zooming in on the front angle, we can consider the problem of Stokes flow around an infinite wedge of opening angle $\oangle$. Exact solutions can be obtained to this problem via separation of variables in polar coordinates $(r, \theta)$ \cite{poz1997}. In this section alone, $\theta$ denotes the polar angle and not the tangent angle of the body.
The corresponding stream function is $\psi = r^{\lambda}f(\theta)$, where $f$ satisfies the 4th order ODE
\begin{equation}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0
\end{equation}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$, where $\thb = \pi - \oangle/2$, along with the condition of odd-symmetry $f(0) = f''(0) = 0$ (corresponding to up-down symmetry of the flow).
The corresponding shear stress evaluated on the surface of the wedge is $\tau = \mu r^{\lambda-2} f''(\thb)$. Thus, the wedge of uniform stress corresponds to $\lambda = 2$. For this value, the solution for $f$ satisfying odd symmetry is
\begin{equation}
f(\theta) = B \sin (2 \theta) + C \theta
\end{equation}
Imposing the no-slip boundary conditions gives the condition
\begin{equation}
\sin(2 \thb) - 2 \thb \cos(2 \thb) = 0
\end{equation}
which has solution $\thb \approx 129$ degrees. This gives an opening angle of $\oangle = 2(\pi - \thb) \approx 102$ degrees.
